---
title: "Alochol vs. Happiness"
author: "Yoonsung Park"
output: html_notebook
---




```{r}
library(tidyverse)
library(mosaic)
```

```{r}
happinessCSV <- read.csv(file = 'output.csv', header=TRUE, fill = T, na.strings = "")
alcoholCSV <- read.csv(file = 'alcohol-consumption-vs-gdp-per-capita.csv')
```

```{r}
head(happinessCSV)
head(alcoholCSV)
```

## DATA CLEANING



```{r}
#happiness data have multiple data combined by each years, so group by each countries and take mean of 2015-2022
happinessCSV %>% 
  group_by(Country) %>%
  summarise(mean(Happiness.Score.1, na.rm=TRUE))
happinessCSV %>% 
  group_by(Country) %>%
  summarise(mean(Happiness.Score, na.rm=TRUE))
happinessCSV %>% 
  group_by(Country) %>%
  summarise(mean(Score, na.rm=TRUE))
happinessCSV %>% 
  group_by(Country) %>%
  filter(year == 2022) %>%
  summarise(mean(Happiness.score, na.rm=TRUE))
```


```{r}
#getting error in Happiness.score column bc there is comma instead period, removing comma and converting data type.
happinessCSV$Happiness.score <- as.numeric(gsub(",",".",happinessCSV$Happiness.score))
happinessCSV %>% 
  filter(year == 2022)
```

```{r}
#calculate mean value of happiness on corresponding feature and save them as subset dataframe + 
#realized there were more different features for country and happiness
y2015_2016 <-
  happinessCSV %>% 
  group_by(Country) %>%
  summarise(happiness_mean = mean(Happiness.Score.1, na.rm=TRUE))
y2017 <-
  happinessCSV %>% 
  group_by(Country) %>%
  summarise(happiness_mean = mean(Happiness.Score, na.rm=TRUE))
y2018_2019 <- 
  happinessCSV %>% 
  group_by(Country = Country.or.region) %>%
  summarise(happiness_mean = mean(Score, na.rm=TRUE))
y2020_2021 <- 
  happinessCSV %>% 
  group_by(Country = Country.name) %>%
  summarise(happiness_mean = mean(Ladder.score, na.rm=TRUE))
y2022 <-
  happinessCSV %>% 
  group_by(Country) %>%
  summarise(happiness_mean = mean(Happiness.score, na.rm=TRUE))
```


```{r}
#combined the subset data
ycombined <- rbind(y2015_2016, y2017, y2018_2019, y2020_2021, y2022)
ycombined <- 
  ycombined %>%
  group_by(Country) %>%
  summarise(happiness = mean(happiness_mean, na.rm=TRUE))
ycombined
```


```{r}
ycombined %>%
  filter(Country == "Lesotho" | Country == "Lesotho*")
#there are multiple variation in names in same country
```


```{R}
#find list of countries that have *
ycombined %>% 
  filter(grepl("\\*", Country, ignore.case = TRUE)) 
```


```{r}
#create for loop through country and remove * character
for(i in 1:length(ycombined$Country)){
  ycombined$Country[i] <- gsub('\\*', '', ycombined$Country[i])
}
```


```{R}
#check if they are removed
ycombined %>% 
  filter(grepl("\\*", Country, ignore.case = TRUE)) 
```


```{r}
#calculate mean value again
ycombined <- 
  ycombined %>%
  group_by(Country) %>%
  summarise(happiness = mean(happiness, na.rm=TRUE))
ycombined
```


```{r}
#Data cleaning for alcohol data
alcohol <-
  alcoholCSV %>%
  group_by(Country = Entity) %>% 
  summarise(alcohol = mean(Total.alcohol.consumption.per.capita..liters.of.pure.alcohol..projected.estimates..15..years.of.age., na.rm=TRUE))
alcohol
```


```{r}
#Combined happiness and alocohol data
combined <-
  ycombined %>%
  inner_join(alcohol)
combined
```

## Data Visualization

```{r}
#linear regression
ggplot(data = combined, aes(x = alcohol, y = happiness)) + 
  geom_point() + 
  stat_smooth(method = lm) + 
  labs(title = "")

#smooth regression
ggplot(data = combined, aes(x = alcohol, y = happiness)) + 
  geom_point() + 
  stat_smooth(method = loess) + 
  labs(title = "")
```

So, based on the graph we created for the combined table, alcohol consumption does have a positive correlation with happiness scores. 
When alcohol consumption in a country is high, the happiness score seems to be higher, 
when alcohol consumption is low in a country the happiness score seems to be low as well. 
However, the correlation seems to be less obvious as alcohol consumption goes over 12 alcohol level on smooth regression.


```{r}
#see which countries have highest/lowest happiness and alochol level
combined %>%
  arrange(desc(happiness))

combined %>%
  arrange(desc(alcohol))

combined %>%
  arrange(happiness)

combined %>%
  arrange(alcohol)
```

By looking at the countries that have the highest/lowest happiness and alcohol level data table, 
we can clearly see there is some difference in happiness between lowest alcohol consumption versus highest. 
It is important to realize this is not a cause and effect relation but only a correlation. To go on further, 
the cause for correlation might be lower alcohol consumption countries have tighter regulations which possibly can have negative affect on general happiness. 
If we look at the countries that have low alcohol consumption, 
some of the countries have war going on and have religious or government regulation on alcohol. 
On the other hand, Nordic countries are known for being happy and their alcohol culture, affecting the correlation greatly.
